 

### 2021-06-07 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Further, @xuSeasonalVariationTemperature2006 showed that the response of respiration to $T_{leaf}$ [partially represented  by $E_{0}$] in Quercus rubra hardly varied with canopy position. However, considering site differences, canopy-position effect in respiration was much greater in mesic sites, and seasonal effect is greater in drier sites.

@@

Biophysically, greater $VPD$, atmospheric demand and low RH can be more stressful for trees, than soil moisture across biomes-including mesic forests,  as the former limits surface conductance and evapotranspiration to a greater extent than soil moisture [@novickIncreasingImportanceAtmospheric2016; @corlettImpactsWarmingTropical2011; @ruehrWaterAvailabilityDominant2016]. Accordingly, heat, associated with increase in temperature and VPD even under adequete water can cause greater tree hydraulic stress and altered and soil-water dynamics than milder drought, particulary in temperate species [@ruehrWaterAvailabilityDominant2016]. However, soil water loss during drought effects larger trees more than small trees, because limited soil-water cannot meet the evaporative demand via the long-distance water transport to the canopy [Norway spruce study, bennett; @mcgregor_tree_nodate]. Further, heat with drought exacerbates tree stress [@ruehrWaterAvailabilityDominant2016]. Whereas, greater $RH$ can do the opposite, where it stifles evapotranspiration thereby increasing $T_{leaf}$, particularly in tropical canopies where irradiance can have a greater effect [Fig.3; @perezIncreasingHumidityThreatens2018; @tibbittsHumidityPlants1979].

@@

>>>>

Further, @xuSeasonalVariationTemperature2006 showed that the response of respiration to $T_{leaf}$ [partially represented  by $E_{0}$] in Quercus rubra also hardly varied with canopy position. However, considering site differences, canopy-position effect in respiration was much greater in mesic sites, and seasonal effect was greater in drier sites.

@@

Biophysically, greater $VPD$, atmospheric demand and low RH can be more stressful for trees, than soil moisture across biomes-including mesic forests,  as the former limits surface conductance and evapotranspiration to a greater extent than soil moisture [@novickIncreasingImportanceAtmospheric2016; @corlettImpactsWarmingTropical2011; @ruehrWaterAvailabilityDominant2016]. Accordingly, heat, associated with increase in temperature and VPD even under adequete water can cause greater tree hydraulic stress and altered soil-water dynamics than milder drought, particulary in temperate species [@ruehrWaterAvailabilityDominant2016]. However, soil water loss during drought effects larger trees more than small trees, because limited soil-water cannot meet the evaporative demand via the long-distance water transport to the canopy [Norway spruce study, bennett; @mcgregor_tree_nodate]. Further, heat with drought exacerbates tree stress [@ruehrWaterAvailabilityDominant2016]. Whereas, greater $RH$ can do the opposite, where it stifles evapotranspiration thereby increasing $T_{leaf}$, particularly in tropical canopies where irradiance can have a greater effect [Fig.3; @perezIncreasingHumidityThreatens2018; @tibbittsHumidityPlants1979].

@@

-Shading by the overstory can be better represented to predict understory temperatures by modeling the time-varying interaction of the forest canopy with solar radiation, as has been done in other LIDAR-based analyses (e.g., Bode et al. https://doi.org/10.1016/j.rse.2014.01.028, Musselman et al. https://doi.org/10.1016/j.rse.2013.05.021, Davis et al. https://doi.org/10.1016/j.agrformet.2019.02.015)
